---
title: "20231013_SoupX_ambientfiltering_manual"
author: "Jolet Mimpen"
date: "2023-10-13"
output: html_document
---

#Setup
```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/")

getwd()
```


#Load packages
```{r}
library(SoupX)
library(celda)
library(Seurat)
library(SingleCellExperiment)
library(stringr)
library(dplyr)
library(scuttle)
library(ggplot2)

```

#Set up new folder
```{r}
working_directory <- getwd()
working_directory

date <- Sys.Date() %>% str_replace_all("-", "")

dir.create(paste0(working_directory, "/", date, "_decontX_SoupX_scDblFinder/"))
primary_directory <- paste0(working_directory, "/", date, "_decontX_SoupX_scDblFinder/")

dir.create(paste0(primary_directory, "/", "RDS_objects/"))
save_directory <- paste0(primary_directory, "/", "RDS_objects/")

```


#Select sample locations
```{r}
cells_file_list <- list.files("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20230926_14-55_QC-Filter.dir/RDS_objects.dir/filtered/", pattern="_SingleCellExp.rds")

droplets_file_list <- list.files("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20230926_14-55_QC-Filter.dir/RDS_objects.dir/raw/", pattern="_SingleCellExp.rds")


```

#Choose files to open
```{r}

file_number <- 1

cells_sce <- readRDS(paste0("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20230926_14-55_QC-Filter.dir/RDS_objects.dir/filtered/", cells_file_list[file_number]))

droplets_sce <- readRDS(paste0("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20230926_14-55_QC-Filter.dir/RDS_objects.dir/raw/", droplets_file_list[file_number]))

cells_sce@colData$sample[1]
droplets_sce@colData$sample[1]

current_sample <- cells_sce@colData$sample[1]

dir.create(paste0(primary_directory, current_sample, "/"))

sub_directory <- paste0(primary_directory, current_sample, "/")
sub_directory
```

#Get list of matched ensembl and gene names
```{r}
library(biomaRt)

ensembl <- try(useMart("ensembl", dataset = "hsapiens_gene_ensembl"))
if(class(ensembl) == "try-error"){
  httr::set_config(httr::config(ssl_verifypeer = FALSE))
  ensembl <- useMart("ensembl", dataset = "hsapiens_gene_ensembl")
}

# get a df of ensemble_id vs gene symbol
mapping <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), mart = ensembl)
```


#Add gene names to droplets object using BioMart
```{r, echo=FALSE, message = FALSE}


#make a df with ensemble_id and hgnc_symbol
df <- data.frame("ensembl_gene_id" = rownames(droplets_sce)) %>% 
        left_join(mapping)
    
    #add a column of unique names, remove the "_" from ensembl_ids
    df$unique_name <- uniquifyFeatureNames(df$ensembl_gene_id, df$hgnc_symbol)
    df$unique_name <- gsub("^_", "", df$unique_name)
    df <- df %>% distinct(ensembl_gene_id, .keep_all = TRUE) #remove duplicated rows
    
    #add ensembl_IDs to row names
    rownames(df) <- rownames(droplets_sce)
    
    #add to the meta.features in the RNA assay
    rowData(droplets_sce) <- df
    
    # Update the assay rownames
    rownames(droplets_sce) <- rowData(droplets_sce)$unique_name
    
    rownames(cells_sce) <- rowData(cells_sce)$unique_name


```


#OR Add gene name by copying across
```{r}
rownames(droplets_so) <- rownames(cells_so)
```


#Convert to Seurat
```{r}
    
#Convert to Seurat object
droplets_so <- CreateSeuratObject(counts = counts(droplets_sce),
                                     assay = "RNA",
                                     project = mainExpName(droplets_sce),
                                     meta.data = as.data.frame(colData(droplets_sce)))
```





#Convert cells SCE into Seurat Object, normalise/scale/etc, plot
```{r}
cells_so <- CreateSeuratObject(counts = counts(cells_sce),
                                     assay = "RNA",
                                     project = mainExpName(cells_sce),
                                     meta.data = as.data.frame(colData(cells_sce)))

vlnplot1 <- VlnPlot(cells_so, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2, log = T)
vlnplot2 <- VlnPlot(cells_so, features = c("subsets_mito_percent"), ncol = 1)
(vlnplot1 | vlnplot2)

cells_so <- cells_so %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData() %>%
           RunPCA()%>%
           FindNeighbors(dims = 1:50) %>%
           RunUMAP(dims = 1:50) %>%
           FindClusters(resolution = 0.3)



DimPlot(cells_so, reduction = "umap", label = T)

```


#Add Seurat cluster info into SCE
```{r}

# put the cluster info in the sce colData
colData(cells_sce)$seurat_clusters <- cells_so[["seurat_clusters"]]
# Add the PCA & UMAP dim reductions to the sce object
reducedDim(cells_sce) <- Embeddings(cells_so, reduction = "pca")
reducedDim(cells_sce, type = "umap") <- Embeddings(cells_so, reduction = "umap")
    

cells_sce <- celda::decontX(cells_sce, 
                            background = droplets_sce, 
                            z = pull(colData(cells_sce)$seurat_clusters))


plotDecontXContamination(cells_sce)


```



#Convert cells SCE into Seurat Object, normalise/scale/etc, plot
```{r}

cells_so <- CreateSeuratObject(counts = counts(cells_sce),
                                     assay = "RNA",
                                     project = mainExpName(cells_sce),
                                     meta.data = as.data.frame(colData(cells_sce)))

cells_so[["decontXcounts"]] <- CreateAssayObject(counts = decontXcounts(cells_sce))

DefaultAssay(cells_so) <- "RNA"

cells_so <- cells_so %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData() %>%
           RunPCA()%>%
           FindNeighbors(dims = 1:50) %>%
           RunUMAP(dims = 1:50) %>%
           FindClusters(resolution = 0.3)


vlnplot1 <- VlnPlot(cells_so, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2, log = T)
vlnplot2 <- VlnPlot(cells_so, features = c("subsets_mito_percent", "decontX_contamination"), ncol = 2)
vlnplot1
vlnplot2

umapplot <- DimPlot(cells_so, reduction = "umap", label = T)
umapplot

decontXcontaminationplot <- FeaturePlot(cells_so, 
                                features = "decontX_contamination",
                                cols = c("grey89", "red"))

decontXcontaminationplot

rm(vlnplot1, vlnplot2)

```


#Save UMAPs
```{r}

ggsave(paste0(sub_directory, "/UMAP_RNA.png"), plot = umapplot, width = 8, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/UMAP_decontXcontamination.png"), plot = decontXcontaminationplot, width = 8, height = 5, bg = "white")

rm(umapplot, decontXcontaminationplot)
```



#Extract counts matrices from cells and droplet Seurat object and reorder rownames to ensure that they are the same order
```{r}

library(Matrix)

cell.matrix <- cells_so@assays$RNA@counts
cell.matrix <- Matrix(cell.matrix, sparse = T)
cell.matrix <- cell.matrix[order(row.names(cell.matrix)), ]


droplet.matrix <- droplets_so@assays$RNA@counts
droplet.matrix <- Matrix(droplet.matrix, sparse = T)
droplet.matrix <- droplet.matrix[order(row.names(droplet.matrix)), ]

identical(rownames(cell.matrix), rownames(droplet.matrix))

```


#Make Soup Channel
```{r}
soup.channel <- SoupChannel(droplet.matrix, cell.matrix)
```


#Add metadata and UMAP to Soup Channel
```{r}

cell.metadata <- cells_so[[]]
cell.umap <- cells_so@reductions$umap@cell.embeddings

soup.channel  <- setClusters(soup.channel, setNames(cell.metadata$seurat_clusters, rownames(cell.metadata)))
soup.channel  <- setDR(soup.channel, cell.umap)
```


#Adipocyte markers
```{r, fig.height=5, fig.width=10}
adipocytemarkers <- c("ADIPOQ", "PLIN1", "PLIN4", "GPAM", "AQP7")

DimPlot(cells_so, reduction = "umap", label = T)
FeaturePlot(cells_so, features = adipocytemarkers, ncol = 3)
DotPlot(cells_so, features = adipocytemarkers, cols = c("blue", "red"))
VlnPlot(cells_so, features = adipocytemarkers)
```

#VEC markers
```{r, fig.height=5, fig.width=10}
VECmarkers <- c("VWF", "PTPRB", "FLT1")

DimPlot(cells_so, reduction = "umap", label = T)
FeaturePlot(cells_so, features = VECmarkers, ncol = 3)
DotPlot(cells_so, features = VECmarkers, cols = c("blue", "red"))
VlnPlot(cells_so, features = VECmarkers)
```

#Macrophage markers
```{r, fig.height=5, fig.width=10}
macrophagemarkers <- c("CD163", "CD163L1", "CD14", "CD68", "MRC1", "MSR1")

DimPlot(cells_so, reduction = "umap", label = T)
FeaturePlot(cells_so, features = macrophagemarkers, ncol = 3)
DotPlot(cells_so, features = macrophagemarkers, cols = c("blue", "red"))
VlnPlot(cells_so, features = macrophagemarkers)
```

#Fibroblast markers
```{r, fig.height=5, fig.width=10}
fibroblastmarkers <- c("COL1A1", "COL1A2", "COL3A1", "DCN")

DimPlot(cells_so, reduction = "umap", label = T)
FeaturePlot(cells_so, features = fibroblastmarkers, ncol = 4)
DotPlot(cells_so, features = fibroblastmarkers, cols = c("blue", "red"))
VlnPlot(cells_so, features = fibroblastmarkers, ncol = 4)

FeaturePlot(cells_so, features = "NOTCH3") | DotPlot(cells_so, features = "NOTCH3", cols = c("blue", "red")) | VlnPlot(cells_so, features = "NOTCH3")
```


#Head the top 20 most highly expressed genes in the background
```{r}
head(soup.channel$soupProfile[order(soup.channel$soupProfile$est, decreasing = TRUE), ], n = 20)

```


#Estimate which cells should not express the list of markers. Play around with the maximum contamination number to get the right cells (which you can do based on expression of these markers on the feature plots above)
##Lining layer markers
```{r}
markerlisttouse <- fibroblastmarkers
markerlistname <- "fibroblastmarkers"

useToEst = estimateNonExpressingCells(soup.channel, nonExpressedGeneList = list(markerlisttouse), maximumContamination = 8) 


markermapplot <- plotMarkerMap(soup.channel, geneSet = markerlisttouse, useToEst = useToEst)
markermapplot
```


#Calculate contamination fraction
```{r}
soup.channel <- calculateContaminationFraction(soup.channel, list(markerlisttouse), useToEst = useToEst, forceAccept = T)
```



#Export matrix of adjusted counts
```{r}
out <- adjustCounts(soup.channel)
```


#Check if counts were appropriately adjusted
```{r}
PRG4plot <- plotChangeMap(soup.channel, out, "PRG4") + ggtitle("PRG4")
PRG4plot
FN1plot <- plotChangeMap(soup.channel, out, "FN1") + ggtitle("FN1")
FN1plot
COL1A1plot <- plotChangeMap(soup.channel, out, "COL1A1") + ggtitle("COL1A1")
COL1A1plot
COL1A2plot <- plotChangeMap(soup.channel, out, "COL1A2") + ggtitle("COL1A2")
COL1A2plot
COL3A1plot <- plotChangeMap(soup.channel, out, "COL3A1") + ggtitle("COL3A1")
COL3A1plot
KAZNplot <- plotChangeMap(soup.channel, out, "KAZN") + ggtitle("KAZN")
KAZNplot
CLUplot <- plotChangeMap(soup.channel, out, "CLU") + ggtitle("CLU")
CLUplot
DCNplot <- plotChangeMap(soup.channel, out, "DCN") + ggtitle("DCN")
DCNplot

KITplot <- plotChangeMap(soup.channel, out, "KIT") + ggtitle("KIT")
KITplot

PLIN1plot <- plotChangeMap(soup.channel, out, "PLIN1") + ggtitle("PLIN1")
PLIN1plot

CD163plot <- plotChangeMap(soup.channel, out, "CD163") + ggtitle("CD163")
CD163plot

plotChangeMap(soup.channel, out, "APOE") + ggtitle("APOE")
VWFplot <- plotChangeMap(soup.channel, out, "VWF") + ggtitle("VWF")
VWFplot

THEMISplot <- plotChangeMap(soup.channel, out, "THEMIS") + ggtitle("THEMIS")
THEMISplot
MS4A1plot <- plotChangeMap(soup.channel, out, "MS4A1") + ggtitle("MS4A1")
MS4A1plot

NOTCH3plot <- plotChangeMap(soup.channel, out, "NOTCH3") + ggtitle("NOTCH3")
NOTCH3plot

HLADRAplot <- plotChangeMap(soup.channel, out, "HLA-DRA") + ggtitle("HLA-DRA")
HLADRAplot

TTNplot <- plotChangeMap(soup.channel, out, "TTN") + ggtitle("TTN")
TTNplot

TRDNplot <- plotChangeMap(soup.channel, out, "TRDN") + ggtitle("TRDN")
TRDNplot
```

```{r, fig.height=15, fig.width=20}
plotchangemapgrid <- cowplot::plot_grid(PRG4plot, FN1plot, DCNplot, CLUplot, COL1A1plot, COL1A2plot, COL3A1plot, KAZNplot, CD163plot, THEMISplot, MS4A1plot, KITplot, VWFplot, NOTCH3plot, PLIN1plot, TTNplot, ncol = 4)

plotchangemapgrid
```


#Save UMAPs
```{r}

ggsave(paste0(sub_directory, "/UMAP_plotMarkerMap_", markerlistname, ".png"), plot = markermapplot, width = 8, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/UMAP_plotchangemapgrid_", markerlistname, ".png"), plot = plotchangemapgrid, width = 15, height = 20, bg = "white")

rm(PRG4plot, FN1plot, DCNplot, CLUplot, COL1A1plot, COL1A2plot, COL3A1plot, KAZNplot, HLADRAplot, CD163plot, THEMISplot, MS4A1plot, KITplot, VWFplot, NOTCH3plot, PLIN1plot, TTNplot)

```


#Add "out"-channel into your Seurat object
```{r}

new_so <- cells_so

new_so[["SoupXcounts"]] <- CreateAssayObject(counts = out)

DimPlot(new_so, reduction = "umap", label = T)

#saveRDS(new_so, "/ceph/project/tendonhca/shared/chromium/analysis/20230302_synovium/202310_SoupX_adjusted/MSK0787-SYN-KNEE_filtered_SoupX_removed")


```



#Normalise and scale soupX counts so that they can be used for plotting
```{r}
DefaultAssay(object = new_so) <- "decontXcounts"

new_so <- new_so %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData()

DefaultAssay(object = new_so) <- "SoupXcounts"

new_so <- new_so %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData()


DimPlot(new_so, reduction = "umap", label = T)


```

#Make featureplots of list of genes with RNA and SoupXcounts matrices
```{r, fig.height=15, fig.width=15}
#genelist <- c("PRG4", "CLIC5", "CD55", "HBEGF", "FN1", "CRTAC1", "ITGB8", "CLU", "PLIN1", "COL1A1", "COL1A2", "COL3A1", "DCLK1", "IGHA1", "HLA-DRA", "APOE")

genelist <- c("COL1A1", "COL1A2", "COL3A1", "DCLK1", "PRG4", "FN1", "TTN", "TRDN", "CD163", "PLIN1", "VWF", "THEMIS", "MS4A1", "NOTCH3")

DefaultAssay(object = new_so) <- "RNA"
plot1 <- FeaturePlot(new_so, features = genelist)


DefaultAssay(object = new_so) <- "decontXcounts"
plot2 <- FeaturePlot(new_so, features = genelist)


DefaultAssay(object = new_so) <- "SoupXcounts"
plot3 <- FeaturePlot(new_so, features = genelist)



plot1
plot2
plot3

```


```{r, fig.height=15, fig.width=45}
plot0 <- cowplot::plot_grid(plot1, plot2, plot3, ncol = 3)
plot0
```



#Save Featureplots
```{r}

ggsave(paste0(sub_directory, "/Featureplots_RNA.png"), plot = plot1, width = 15, height = 15, bg = "white")
ggsave(paste0(sub_directory, "/Featureplots_decontXcounts.png"), plot = plot2, width = 15, height = 15, bg = "white")
ggsave(paste0(sub_directory, "/Featureplots_SoupXcounts.png"), plot = plot3, width = 15, height = 15, bg = "white")
ggsave(paste0(sub_directory, "/Featureplots_combined.png"), plot = plot0, width = 45, height = 15, bg = "white")

rm(plot1, plot2, plot3, plot0)
```



#Make dotplots of list of genes with RNA and SoupXcounts matrices
```{r, fig.width=12, fig.height=5}

geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN", "PRG4", "CDH11", "FN1", "POSTN", 
              "PECAM1", "PTPRB", "FLT1", "VWF", "VEGFC",
              "NOTCH3", "PDGFRB", "MYO1B", 
              "MMRN1", "PROX1", "KDR", "FLT4", 
              "GPAM", "AQP7", "ADIPOQ", 
              "PTPRC", "CD247", 
              "CD69", "KIT", "CDK15",
              "MS4A1", "CD37", "BLNK", "SDC1", "IGHA1", 
               "CD163", "MERTK", "MRC1",
              "ASPM", "DIAPH3",
              "IRF7", "CLEC4C", "IL3RA", "HLA-DQA1",
              "TTN", "TRDN", "NEB", 
              "PAX7")

DefaultAssay(object = new_so) <- "RNA"
plot4 <- DotPlot(new_so, features = geneName, cols = c("blue", "red"), assay = "RNA")

DefaultAssay(object = new_so) <- "decontXcounts"
plot5 <- DotPlot(new_so, features = geneName, cols = c("blue", "red"), assay = "decontXcounts")

DefaultAssay(object = new_so) <- "SoupXcounts"
plot6 <- DotPlot(new_so, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts")



plot4 <- plot4 + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("RNA")
plot5 <- plot5 + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("decontXcounts")
plot6 <- plot6 + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("SoupXcounts")

plot4
plot5
plot6


```


#Dotplot grid to compare
```{r, fig.width=12, fig.height=15}
plot7 <- cowplot::plot_grid(plot4, plot5, plot6, ncol = 1)

plot7
```


#Save dotplots
```{r}

ggsave(paste0(sub_directory, "/Dotplot_RNA.png"), plot = plot4, width = 12, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/Dotplot_decontXcounts.png"), plot = plot5, width = 12, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/Dotplot_SoupXcounts.png"), plot = plot6, width = 12, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/Dotplot_combined.png"), plot = plot7, width = 12, height = 15, bg = "white")

rm(plot4, plot5, plot6, plot7)
```


#Run scDblFinder and add metadata to Seurat object
```{r}
library(scDblFinder)

new_so_scdblmetrics <- scDblFinder(new_so@assays$SoupXcounts@counts, clusters = new_so@meta.data$seurat_clusters)


new_so@meta.data$scDblFinder.class <- new_so_scdblmetrics@colData$scDblFinder.class
new_so@meta.data$scDblFinder.score <- new_so_scdblmetrics@colData$scDblFinder.score

```


#Check new clustering
```{r, fig.height=6, fig.width=10}
DefaultAssay(object = new_so) <- "SoupXcounts"

new_so <- new_so %>% 
           NormalizeData() %>%
           FindVariableFeatures() %>% 
           ScaleData() %>%
           RunPCA()%>%
           FindNeighbors(dims = 1:50) %>%
           RunUMAP(dims = 1:50) %>%
           FindClusters(resolution = 0.3)

newumap <- DimPlot(new_so, reduction = "umap", label = T)
newumap_scDbl <- DimPlot(new_so, reduction = "umap", label = T, group.by = "scDblFinder.class")
newumap_decontX <- FeaturePlot(new_so, reduction = "umap", features = "decontX_contamination")
plot5 <- DotPlot(new_so, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts")
newdotplot <- plot5 + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

newumap
newumap_scDbl
newumap_decontX
newdotplot

```


#Save plots
```{r}

ggsave(paste0(sub_directory, "/SoupXclustered_UMAP.png"), plot = newumap, width = 8, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/SoupXclustered_UMAP_decontXcontamination.png"), plot = newumap_decontX, width = 8, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/SoupXclustered_UMAP_scDblFinder.png"), plot = newumap_scDbl, width = 8, height = 5, bg = "white")
ggsave(paste0(sub_directory, "/SoupXclustered_UMAP_Dotplot.png"), plot = newdotplot, width = 12, height = 5, bg = "white")

rm(newumap, newumap_decontX, newumap_scDbl, newdotplot)

```


#Save RDS
```{r}

saveRDS(new_so, paste0(save_directory, current_sample, "_", markerlistname, "_filtered_decontX_SoupX_scDblFinder.RDS"))

```




#Remove 
```{r}
rm(cell.matrix, cell.metadata, cell.umap, cells_sce, cells_so, df, droplet.matrix, droplets_sce, droplets_so, ensembl, mapping, new_so, new_so_scdblmetrics, out, soup.channel, useToEst)

rm(adipocytemarkers, cells_file_list, current_sample, date, droplets_file_list, fibroblastmarkers, macrophagemarkers, markerlistname, markerlisttouse, VECmarkers, save_directory, sub_directory, working_directory, primary_directory)

rm(decontXcontaminationplot, df, markermapplot, plot5, plotchangemapgrid, TRDNplot, umapplot)
```


