---
title: "20240409_quads_fibroblastsonly"
author: "Jolet Mimpen"
date: "2024-04-09"
output: html_document
---


```{r setup, load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
library(Matrix)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(harmony)
library(cowplot)
library(clustree)
library(sctransform)
#library(scCustomize)
library(scuttle)
library(celda)

```


#Data retrieval 
```{r}
quads_data <- readRDS("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20231130_quads_int_labelled.RDS")

```



#Subset Stromal only
```{r, fig.width=12, fig.height=4}

#Choose one of the two options below:
quads_fibroblasts <- subset(quads_data, idents = c("Fibroblasts", "Dividing fibroblasts / mural cells"))

#Then continue with Seurat/Harmony
quads_fibroblasts <- quads_fibroblasts %>% 
                Seurat::NormalizeData(verbose = FALSE) %>%
                FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
                ScaleData(verbose=FALSE) %>% 
                RunPCA(npcs = 50, verbose=FALSE)

quads_fibroblasts <- RunHarmony(quads_fibroblasts, group.by.vars = "orig.ident", plot_convergence = TRUE)

ElbowPlot(quads_fibroblasts, ndims = 50)

```

#Make clustree
```{r}

library(clustree)

dims.choice <- 20
resolution.range <- seq(from = 0, to = 1.0, by = 0.1)
quads_fibroblasts <- FindNeighbors(quads_fibroblasts, reduction = "harmony", dims = 1:dims.choice)
quads_fibroblasts <- FindClusters(quads_fibroblasts, resolution = resolution.range)

clustree(quads_fibroblasts, prefix = "SoupXcounts_snn_res.")


```



#UMAPs for every resolution
```{r, fig.width=9, fig.height=5}

quads_fibroblasts <- RunUMAP(quads_fibroblasts, reduction = "harmony", dims = 1:dims.choice)

for (i in resolution.range){
  Idents(quads_fibroblasts) <- paste0("SoupXcounts_snn_res.", i)
  print(DimPlot(quads_fibroblasts, reduction = "umap", label = T) + ggtitle(paste0("Resolution ", i)) + theme(plot.title = element_text(hjust = 0.5)))
}

```


#Choose resolution and plot UMAP
```{r, fig.width=12, fig.height=6}

Idents(quads_fibroblasts) <- "SoupXcounts_snn_res.0.2"

DimPlot(quads_fibroblasts, reduction = "umap", label = F, repel = T, split.by = "tendon_disease")
DimPlot(quads_fibroblasts, reduction = "umap", label = TRUE, repel = T, split.by = "tendon_disease")
```



#Plot Stromal only - check key markers
```{r, fig.width=12, fig.height=6}
geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN",
              "NEGR1", "HMCN1", "DCLK1",
              "CLU", "COMP",  "CILP", "NOX4", "PRG4",
              "PTPRC", "CD163", "MRC1", "MSR1", "MERTK", "LYVE1", "CLEC10A",
              "PECAM1", "PTPRB", "FLT1", "VWF",
              "TRDN", "TTN", "NEB", "TNNT1", "TNNT3", "PAX7",
              "NOTCH3", "PDGFRB", "MYO1B",
              "PLIN1", "AQP7", "ADIPOQ", 
              "CD247", "SKAP1", "THEMIS",
              "MMRN1", "PROX1", "PKHD1L1",
              "ASPM", "DIAPH3", "TOP2A",
              "BCL11A", "CUX2", "CLEC4C",
              "IBSP", "SP7", "RUNX2",
               "KIT", "CPA3", "IL18R1",
              "MMP9", "ACP5", "SIGLEC15",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2")


DotPlot(quads_fibroblasts, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - all stromal")

DotPlot(quads_fibroblasts, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts", split.by = "tendon_disease") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - all stromal")


geneName <- c("COL1A1", "COL1A2", "COL3A1",  "DCN", "PRG4",
              "ABCA8", "COL15A1", "CD55", "DCLK1", "ELN", "FBN1", "FBLN1", "FBLN2", "FBLN5", "KCND2", "NEGR1", "NOVA1", "PDGFRA", "VIT",
              "CADM1", "COL11A1", "COL11A2", "COL12A1", "COL24A1", "COMP", "CPXM2", "FMOD", "MET", "MKX", "ITGA10", "PIEZO2", "THBS4", "THSD4", "TNMD")

DotPlot(quads_fibroblasts, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

DotPlot(quads_fibroblasts, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts", split.by = "tendon_disease") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - all stromal")



```


#Make Heatmap stromal only only
```{r, fig.width=8, fig.height=10}
stromal.submarkers <- FindAllMarkers(quads_fibroblasts, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

stromal.submarkers %>%
    group_by(cluster) %>%
    top_n(n = 25, wt = avg_log2FC) -> stromal.submarkers.top25


DoHeatmap(quads_fibroblasts, features = stromal.submarkers.top25$gene) + scale_fill_viridis() + NoLegend()



stromal.submarkers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> stromal.submarkers.top10


DoHeatmap(quads_fibroblasts, features = stromal.submarkers.top10$gene) + scale_fill_viridis() + NoLegend()

```


#DotPlot with top10 markers
```{r, fig.width=20, fig.height=6}
geneName <- stromal.submarkers.top10$gene
geneName <- unique(geneName)

DotPlot(quads_fibroblasts, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - all stromal - res0.2")


#VEC4vsVEC6 <- FindMarkers(quads_fibroblasts, ident.1 = "4", ident.2 = c("6"), verbose = T)
#VEC4vsVEC6$geneName <- rownames(VEC4vsVEC6)

```



#Add labels for each cluster of overall object
```{r, fig.height=6, fig.width=10}

new.cluster.ids <- c("FBLNhi fibroblasts", "COL3A1hi fibroblasts", "ABCA10hi fibroblasts", "NR4A1hi fibroblasts", "Dividing cells")

names(new.cluster.ids) <- levels(quads_fibroblasts)
quads_fibroblasts <- RenameIdents(quads_fibroblasts, new.cluster.ids)


DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = F, label.size = 5.0, pt.size = 0.8) + NoLegend()

DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = TRUE) + NoLegend()

DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = TRUE, split.by = "tendon_disease") + NoLegend()


#add names into the metadata
quads_fibroblasts@meta.data$fibroblastcluster_id <- Idents(quads_fibroblasts)

```


#Make DimPlot with labelled cells and new colours
```{r, fig.width=12, fig.height=5}
DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, split.by = "tendon_disease", cols = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + NoLegend()

DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, cols = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + NoLegend()


umap_stromalonly_nolabel <- DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = F, label.size = 5.0, split.by = "tendon_disease", cols = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1"))
umap_stromalonly_nolabel

ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202404_figures/Stromalonly_umap_nolabel.png", plot = umap_stromalonly_nolabel, width = 12, height = 5, bg = "white")

rm(umap_stromalonly_nolabel)

```



```{r, fig.height=4, fig.width=6}
DimPlot(quads_fibroblasts, reduction = "umap", repel = TRUE, label = F, label.size = 5.0, cols = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1"), ) + NoLegend()

DimPlot(quads_fibroblasts, reduction = "umap", group.by = "tendon_disease", cols = c("#eb7323", "#4b2480")) + ggtitle(NULL) + NoLegend()

```



#Plot heatmap of DEGs
```{r, fig.height=9.5, fig.width=6}
DoHeatmap(quads_fibroblasts, features = stromal.submarkers.top10$gene, label = F, group.colors = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + scale_fill_viridis() + NoLegend()

DoHeatmap(quads_fibroblasts, features = stromal.submarkers.top10$gene, label = F, group.colors = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + scale_fill_viridis()

```


#Violin plots for collagens
```{r, fig.height=4.2, fig.width=5}
VlnPlot(quads_fibroblasts, features = c("COL1A1"), cols = c("#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + NoLegend()
VlnPlot(quads_fibroblasts, features = c("COL1A2"), cols = c("#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + NoLegend()
VlnPlot(quads_fibroblasts, features = c("COL3A1"), cols = c("#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1")) + NoLegend()

```



#Plot UMAPs with annotation and new colours
```{r, fig.height=5, fig.width=12}
DimPlot(quads_fibroblasts, split.by = "tendon_disease", cols = c( "#99cc00", "#FFA319FF", "#B09C85FF" , "#CC3333", "#a1caf1"))
```


#FeaturePlots for highly DEGs
```{r, fig.height=4, fig.width=15}
FeaturePlot(quads_fibroblasts, features = c("FBLN1", "NOX4", "CILP"), ncol = 3)
FeaturePlot(quads_fibroblasts, features = c("ADAM12", "COL3A1", "TNC"), ncol = 3)
FeaturePlot(quads_fibroblasts, features = c("ABCA10", "CNTN4", "ABCA8"), ncol = 3)
FeaturePlot(quads_fibroblasts, features = c("NR4A1", "NR4A3", "SEMA4A"), ncol = 3)

```




#FindMarkers for each fibroblast cell type
```{r}
levels(quads_fibroblasts$fibroblastcluster_id)

FBLNhi_fibs_markers <- FindMarkers(quads_fibroblasts, ident.1 = "FBLN1hi fibroblasts", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
COL3A1hi_fibs_markers <- FindMarkers(quads_fibroblasts, ident.1 = "ADAM12hi fibroblasts", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
ABCA10hi_fibs_markers <- FindMarkers(quads_fibroblasts, ident.1 = "ABCA10hi fibroblasts", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
NR4A1hi_fibs_markers <- FindMarkers(quads_fibroblasts, ident.1 = "NR4A1hi fibroblasts", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
Dividing_fibs_markers <- FindMarkers(quads_fibroblasts, ident.1 = "Dividing cells", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

FBLNhi_fibs_markers$geneName <- rownames(FBLNhi_fibs_markers)
COL3A1hi_fibs_markers$geneName <- rownames(COL3A1hi_fibs_markers)
ABCA10hi_fibs_markers$geneName <- rownames(ABCA10hi_fibs_markers)
NR4A1hi_fibs_markers$geneName <- rownames(NR4A1hi_fibs_markers)
Dividing_fibs_markers$geneName <- rownames(Dividing_fibs_markers)

Fibroblast_markers <- list()
Fibroblast_markers[[1]] <- FBLNhi_fibs_markers
Fibroblast_markers[[2]] <- COL3A1hi_fibs_markers
Fibroblast_markers[[3]] <- ABCA10hi_fibs_markers
Fibroblast_markers[[4]] <- NR4A1hi_fibs_markers
Fibroblast_markers[[5]] <- Dividing_fibs_markers
names(Fibroblast_markers) <- c("FBLNhi_fibs_markers", "COL3A1hi_fibs_markers", "ABCA10hi_fibs_markers", "NR4A1hi_fibs_markers", "Dividing_fibs_markers")

```


#Run GO:BP for each fibroblast subset
```{r}

library(gprofiler2)

GOBP_results_fibroblasts <- list()

for(i in 1:length(Fibroblast_markers)){
    
    
set.seed(500)
topcluster <- Fibroblast_markers[[i]] %>% 
  filter(avg_log2FC > 0.5) %>%
  top_n(n=100, wt = avg_log2FC)
gost.res <- gost(topcluster$geneName, organism = "hsapiens", correction_method = "fdr", sources = "GO:BP")

GOBP_results_fibroblasts[[i]] <- gost.res$result
    

      
}

names(GOBP_results_fibroblasts) <- c("FBLNhi_fibs_markers", "COL3A1hi_fibs_markers", "ABCA10hi_fibs_markers", "NR4A1hi_fibs_markers", "Dividing_fibs_markers")

```





#Save RDS object
```{r}
saveRDS(quads_fibroblasts, file = "/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS/202419_quads_fibroblasts_annotated.RDS")

#to re-open quads_fibroblast file:
#quads_fibroblasts <- readRDS("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS/202419_quads_fibroblasts_annotated.RDS")
```

