---
title: "202404_quadsdata_endothelial"
author: "Jolet Mimpen"
date: "2024-04-05"
output: html_document
---



```{r setup, load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
library(Matrix)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(harmony)
library(cowplot)
library(clustree)
library(sctransform)
#library(scCustomize)
library(scuttle)
library(celda)

```


#Data retrieval 
```{r}
quads_data <- readRDS("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20231130_quads_int_labelled.RDS")

```


#Subset endothelial cells
```{r, fig.width=12, fig.height=4}

Idents(quads_data) <- quads_data$cluster_id

#Choose one of the two options below:
quads_endothelial <- subset(quads_data, idents = c("Vascular endothelial cells", "Lymphatic endothelial cells", "Mural cells"))

#Then continue with Seurat/Harmony
quads_endothelial <- quads_endothelial %>% 
                Seurat::NormalizeData(verbose = FALSE) %>%
                FindVariableFeatures(selection.method = "vst", nfeatures = 1000) %>% 
                ScaleData(verbose=FALSE) %>% 
                RunPCA(npcs = 50, verbose=FALSE)

quads_endothelial <- RunHarmony(quads_endothelial, group.by.vars = "orig.ident", plot_convergence = TRUE)

ElbowPlot(quads_endothelial, ndims = 50)



```


#Make clustree
```{r}

library(clustree)

dims.choice <- 30
resolution.range <- seq(from = 0, to = 1.0, by = 0.1)
quads_endothelial <- FindNeighbors(quads_endothelial, reduction = "harmony", dims = 1:dims.choice)
quads_endothelial <- FindClusters(quads_endothelial, resolution = resolution.range)

clustree(quads_endothelial, prefix = "SoupXcounts_snn_res.")


```


#UMAPs for every resolution
```{r, fig.width=9, fig.height=5}

quads_endothelial <- RunUMAP(quads_endothelial, reduction = "harmony", dims = 1:dims.choice)

for (i in resolution.range){
  Idents(quads_endothelial) <- paste0("SoupXcounts_snn_res.", i)
  print(DimPlot(quads_endothelial, reduction = "umap", label = T) + ggtitle(paste0("Resolution ", i)) + theme(plot.title = element_text(hjust = 0.5)))
}

```

#Select resolution
```{r}
Idents(quads_endothelial) <- "SoupXcounts_snn_res.0.5"
```


#DotPlots with known markers
```{r, fig.height=4, fig.width=18}

#Idents(quads_endothelial) <- "endothelial_clusterid"


geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN",
              "NEGR1", "HMCN1", "DCLK1",
              "CLU", "COMP",  "CILP", "NOX4", "PRG4",
              "PTPRC", "CD163", "MRC1", "MSR1", "MERTK", "LYVE1", "CLEC10A",
              "PECAM1", "PTPRB", "FLT1", "VWF",
              "TRDN", "TTN", "NEB", "TNNT1", "TNNT3", "PAX7",
              "NOTCH3", "PDGFRB", "MYO1B",
              "PLIN1", "AQP7", "ADIPOQ", 
              "CD247", "SKAP1", "THEMIS",
              "MMRN1", "PROX1", "PKHD1L1",
              "ASPM", "DIAPH3", "TOP2A",
              "BCL11A", "CUX2", "CLEC4C",
              "IBSP", "SP7", "RUNX2",
               "KIT", "CPA3", "IL18R1",
              "MMP9", "ACP5", "SIGLEC15",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2")


DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) 


geneName <- c("COL1A1", "COL1A2", "COL3A1", "COL3A2",  "DCN", "PRG4",
              "ABCA8", "COL15A1", "CD55", "DCLK1", "ELN", "FBN1", "FBLN1", "FBLN2", "FBLN5", "KCND2", "NEGR1", "NOVA1", "PDGFRA", "VIT",
              "CADM1", "COL11A1", "COL11A2", "COL12A1", "COL24A1", "COMP", "CPXM2", "FMOD", "MET", "MKX", "ITGA10", "PIEZO2", "THBS4", "THSD4", "TNMD")

DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

geneName <- c("COL4A1", "COL4A2", "SPARC", "COL15A1", "VWA", "HSPG2", 
              "CLU", "ACKR1", "CD74", "ADIRF", "HLA-DRA", "TXNIP",
              "JUNB", "FOS", "ZFP36", "JUN", "SOCS3", "JUND",
              "CXCL12", "SLC9A3R2", "SRGN", "GJA4", "PODXL", "EFNB2", 
              "TFPI", "TFF3", "PFIBP1", "EFEMP1", "MMRN", "CCL21") #from Zhang et al. 2023

DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))



geneName <- c("PECAM1", "KDR", "CDH5", "FLT1", "TEK", "CLDN5", "VWF", #panEC genes
              "GJA4", "GJA5", "HEY1", "GATA2", "CXCR4", "SOX17", "MECOM", #artery
              "ACKR1", "PLVAP", "NR2F2", #vein
              "ABCG2", "ABCB1") #VESC (all from https://www.ncbi.nlm.nih.gov/pmc/articles/PMC10070846/)

DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


geneName <- c("FCN3", "EPAS1", "RAMP3", "CDH5", "PODXL", "BTNL9", "NOSTRIN", "FLNB", "CDC42EP1", "CD81", "HPGD", "APP", "EDNRB", "EMCN", "ESAM", "STXBP6", "CLDN5", "CYP3A5", "SOSTDC1", "TBX2", "EPAS1", "SH3BP5", "RALA", "PECAM1", "WWTR1", "GALNT18", "ACVRL1", "SRSF11", "VAT1", "IPO7")
geneName <- unique(geneName)
DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

geneName <- c("IFI27", "TM4SF1", "CLEC14A", "CDH5", "PECAM1", "HYAL2", "SPARCL1", "EGFL7", "CLDN5", "AQP1")
geneName <- c("COL4A1", "COL4A2", "SPARC", "COL15A1", "VWA1", "HSPG2", "PRSS23", "LAMB1", "PLVAP", "APLNR", "IGFBP7", "CLU", "ACKR1", "CD74", "ADIRF", "HLA-DRA", "TXNIP", "JUNB", "FOS", "ZFP36", "JUN", "SOCS3", "JUND", "EIF1", "PPP1R15A", "EGR1", "CXCL12", "SLC9A3R2", "SRGN", "GJA4", "PODXL", "EFNB2", "SRP14", "SEMA3G", "NOTCH4", "HEY1")
geneName <- unique(geneName)
DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


geneName <- c("ACTA2", "TAGLN", "TPM2", "MYH11", "PLN", "SORBS2", "LBH", "IGFBP6", "CNN1", "RERGL", "RAMP1", "RGS5", "APOD", "ABCA9", "STEAP4", "CCL26")
geneName <- unique(geneName)
DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

geneName <- c("RGS5", "ABCC9", "GUCY1A2", "EGFLAM", "FRMD3", "DLC1", "AGT", "PDGFRB", "EPS8", "PLA2G5", "MYH11", "ITGA8", "ACTA2", "TAGLN", "CARMN", "KCNAB1", "ZFHX3", "NTRK3", "PRKG1", "RCAN2")
DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


geneName <- c("PECAM1", "VWF", "PTPRB", "FLT1", "EMCN", #vascular endothelial cells
              "NOTCH3", "PDGFRB", "MYO1B",  "DLC1", "CALD1",  "EBF2", #mural
              "COL6A3", "PDE1C", "COL25A1", "POSTN","STEAP4", "RGS5", #pericytes
              "SLCO2A1", "SELP", "NOSTRIN", "MYRIP", "GNA14", "TACR1", "RAMP3", "LIFR", "ICAM1", #venular
              "PODXL", "EFNB2", "NEBL", "SYT1", "THSD4", "ADAMT16", "CXCL12", "SEMA3G", "SLC9A3R2", "GJA4", "HEY1", #arteriolar
              "COL4A1", "COL4A2",  "COL15A1",  "DYSF",  "ARHGAP18", "LAMB1", "SPARC","GNG2", "VWA1", #capillary
              "PROX1", "MMRN1",  "TFPI",  "RELN", "PKHD1L1", "SEMA3A", "FLRT2", "CCL21", "LYVE1","STAB2", #lymphatic
              "MYH11", "RGS6", "TAGLN", "ADGRL3", "SORBS2", "RCAN2", "MYOCD", "LMOD1", "CARMN", #vSMCs
              "ASPM", "DIAPH3", "TOP2A", "CENPE", "TPX2", "MELK") #dividing
DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```


#Find markers and plot heatmap
```{r, fig.width=8, fig.height=12}
quads_endo_markers <- FindAllMarkers(quads_endothelial, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

quads_endo_markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10.quads.endo

DoHeatmap(quads_endothelial, features = top10.quads.endo$gene) + scale_fill_viridis() + NoLegend()

```


#DotPlot with top10 DEGs in each cluster
```{r, fig.width=20, fig.height=5}
geneName <- top10.quads.endo$gene
geneName <- unique(geneName)

DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
DotPlot(quads_endothelial, features = geneName, cols = c("#eb7323", "#4b2480"), assay = "SoupXcounts", split.by = "tendon_disease") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


```



#Add labels for each cluster of overall object
```{r, fig.height=4, fig.width=6}

Idents(quads_endothelial) <- "SoupXcounts_snn_res.0.5"

new.cluster.ids <- c("Pericytes", "Venular VECs", "Arteriolar VECs", "Capillary VECs", "Lymphatic ECs", "vSMCs", "Dividing VECs")


names(new.cluster.ids) <- levels(quads_endothelial)
quads_endothelial <- RenameIdents(quads_endothelial, new.cluster.ids)


DimPlot(quads_endothelial, reduction = "umap", repel = TRUE, label = F, label.size = 5.0) + NoLegend()

DimPlot(quads_endothelial, reduction = "umap", repel = TRUE, label = TRUE) + NoLegend()

quads_endothelial$endothelial_clusterid <- Idents(quads_endothelial)

```


#Plot DotPlot with the new colours
```{r, fig.width=7, fig.height=5}
DimPlot(quads_endothelial, reduction = "umap", repel = TRUE, label = F, cols = c("#0041c2", "#f04a00", "#b7410e", "#f89880", "#888888", "#4682b4", "#a1caf1")) + NoLegend()

DimPlot(quads_endothelial, reduction = "umap", repel = TRUE, label = F, cols = c("#0041c2", "#f04a00", "#b7410e", "#f89880", "#888888", "#4682b4", "#a1caf1"), split.by = "tendon_disease") 
```



#Plot DotPlot with the new colours
```{r, fig.width=11, fig.height=5}
DimPlot(quads_endothelial, reduction = "umap", repel = TRUE, label = TRUE, cols = c("#0041c2", "#f04a00", "#b7410e", "#f89880", "#888888", "#4682b4", "#a1caf1")) + NoLegend()

DimPlot(quads_endothelial, reduction = "umap", repel = TRUE, label = F, cols = c("#0041c2", "#f04a00", "#b7410e", "#f89880", "#888888", "#4682b4", "#a1caf1"), split.by = "tendon_disease") 
```

#Plot Heatmap with new colours
```{r, fig.width=6, fig.height=8}
DoHeatmap(quads_endothelial, features = top10.quads.endo$gene, label = F, group.colors = c("#0041c2", "#f04a00", "#b7410e", "#f89880", "#888888", "#4682b4", "#a1caf1")) + scale_fill_viridis() + NoLegend()

```

#Make DotPlots with named object
```{r, fig.width=20, fig.height=5}
geneName <- top10.quads.endo$gene
geneName <- unique(geneName)

DotPlot(quads_endothelial, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
DotPlot(quads_endothelial, features = geneName, cols = c("#eb7323", "#4b2480"), assay = "SoupXcounts", split.by = "tendon_disease") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


```

#Proportion table
```{r}
proptable <- prop.table(table(Idents(quads_endothelial), quads_endothelial$patient), margin = 2)
proptable

write.table(proptable, file = "/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202404_figures/proptable.txt", sep = " ")
```


```{r, fig.width=10, fig.height=6}
Idents(quads_endothelial) <- quads_endothelial$cluster_id

DimPlot(quads_endothelial)

Idents(quads_endothelial) <- quads_endothelial$endothelial_clusterid

```



#FindMarkers for each endothelial cell type
```{r}
levels(quads_endothelial$endothelial_clusterid)

Pericytes_markers <- FindMarkers(quads_endothelial, ident.1 = "Pericytes", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
VenularVECs_markers <- FindMarkers(quads_endothelial, ident.1 = "Venular VECs", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
ArteriolarVECs_markers <- FindMarkers(quads_endothelial, ident.1 = "Arteriolar VECs", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
CapillaryVECs_markers <- FindMarkers(quads_endothelial, ident.1 = "Capillary VECs", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
LymphaticECs_markers <- FindMarkers(quads_endothelial, ident.1 = "Lymphatic ECs", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
vSMCs_markers <- FindMarkers(quads_endothelial, ident.1 = "vSMCs", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
DividingVECs_markers <- FindMarkers(quads_endothelial, ident.1 = "Dividing VECs", only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

Pericytes_markers$geneName <- rownames(Pericytes_markers)
VenularVECs_markers$geneName <- rownames(VenularVECs_markers)
ArteriolarVECs_markers$geneName <- rownames(ArteriolarVECs_markers)
CapillaryVECs_markers$geneName <- rownames(CapillaryVECs_markers)
LymphaticECs_markers$geneName <- rownames(LymphaticECs_markers)
vSMCs_markers$geneName <- rownames(vSMCs_markers)
DividingVECs_markers$geneName <- rownames(DividingVECs_markers)

Endothelial_markers <- list()
Endothelial_markers[[1]] <- Pericytes_markers
Endothelial_markers[[2]] <- VenularVECs_markers
Endothelial_markers[[3]] <- ArteriolarVECs_markers
Endothelial_markers[[4]] <- CapillaryVECs_markers
Endothelial_markers[[5]] <- LymphaticECs_markers
Endothelial_markers[[6]] <- vSMCs_markers
Endothelial_markers[[7]] <- DividingVECs_markers

names(Endothelial_markers) <- c("Pericytes_markers", "VenularVECs_markers", "ArteriolarVECs_markers", "CapillaryVECs_markers", "LymphaticECs_markers", "vSMCs_markers", "DividingVECs_markers")

```

#Run GO:BP for each endothelial subset
```{r}

library(gprofiler2)

GOBP_results_endothelial <- list()

for(i in 1:length(Endothelial_markers)){
    
    
set.seed(500)
topcluster <- Endothelial_markers[[i]] %>% 
  filter(avg_log2FC > 0.5) %>%
  top_n(n=100, wt = avg_log2FC)
gost.res <- gost(topcluster$geneName, organism = "hsapiens", correction_method = "fdr", sources = "GO:BP")

GOBP_results_endothelial[[i]] <- gost.res$result
    

      
}

names(GOBP_results_endothelial) <- c("Pericytes_markers", "VenularVECs_markers", "ArteriolarVECs_markers", "CapillaryVECs_markers", "LymphaticECs_markers", "vSMCs_markers", "DividingVECs_markers")

```



#Save RDS file
```{r}
saveRDS(quads_endothelial, file = "/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS/20240408_quads_endothelial_annotated.RDS")

#if need to re-open quads_endothelial:
#quads_endothelial <- readRDS("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS/20240408_quads_endothelial_annotated.RDS")
```

