---
title: "MiloR_Cell_proportions-Quads"
author: "Carla Cohen"
date: "`r Sys.Date()`"
output: html_document
---

# Cell type proportion analysis

Aim: To analyse the how cell proportions vary between healthy and tear in quadriceps tendon

Using MioR as recommended by Bo Sun & Lucy Garner

https://www.nature.com/articles/s41587-021-01033-z  
https://marionilab.github.io/miloR/articles/milo_demo.html  
https://www.bioconductor.org/packages/devel/bioc/vignettes/miloR/inst/doc/milo_gastrulation.html  
https://github.com/MarioniLab/miloR  

## Steps

1. Set up 
- Set up directories, import yml file, set colours  
- Read in Seurat object  

2. Set up the milo object & perform initial calculations
- convert so to sce  
- create a milo object  
- construct KNN graph  
- define representative neighbourhoods & plot  
- compute neighbourhood connectivity  
- count number of cells per neighbourhood  

3.  Set up the experimental design for differential abundance testing  

4.  Perform the differential abundance test  

5. Plot the outputs 
- Inspect the distribution of p values  
- Volcano plot of test results
- Neighbourhood graph on Harmony UMAP projection  
- Beeswarm plot  



### 1. Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(Seurat)
library(tidyverse)
library(miloR)
library(SingleCellExperiment)
library(scater)
library(scran)
library(yaml)
library(cowplot)

set.seed(1) 

# read yaml file

# colours
Okabe_Ito <- c("#117733",  "#ff3d00")
rupture.cols <-  c(Healthy = Okabe_Ito[1], Rupture = Okabe_Ito[2])


quads.colours <- c("#99cc00", "#ff9b19", "#8bc9c9", "#B09C85FF", "#0041c2", "#379975", "#f3ca00", "#f04a00", "#b7410e", "#f89880", "#e37197", "#00a2f3", "#CC3333", "#888888", "#eb0e45", "#4682b4", "#702438", "#09e686", "#a1caf1", "#f0b1d0", "#a1caf1", "#117733", "#a1caf1", "#be0032", "#0a2f8c", "#332288", "#f23f90", "#cf47ed")

names(quads.colours) <- c("FBLN1hi fibroblasts", "ADAM12hi fibroblasts", "MERTKhi LYVE1hi macrophages", "ABCA10hi fibroblasts", "Pericytes", "MERTKhi LYVE1lo macrophages", "Adipocytes", "Venular VECs", "Arteriolar VECs", "Capillary VECs", "T cells", "Nervous system cells", "NR4A1hi fibroblasts", "Lymphatic ECs", "VCANhi DCs/Mono", "vSMCs", "CLEC10Ahi DCs", "MERTKlo PTPRGhi macrophages", "Dividing fibroblasts / mural cells", "NK cells", "Dividing VECs", "Osteoblasts", "Dividing immune cells", "pDCs", "Granulocytes", "Osteoclasts", "B cells", "CLEC9Ahi DCs")

```




Read in Seurat object

```{r, fig.width=10, fig.height=5}
so.quads <- readRDS("/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS/20240524_quads_immune_reannotated.RDS")
so.quads

DimPlot(so.quads, cols = quads.colours)

```

## 2. Set up the milo object & perform initial calculations  
Done following the vignette as I have done in the script miloR_vignette (on computer not cluster)

```{r}

# Convert to single cell experiment object  
sce <- as.SingleCellExperiment(so.quads)

# Create a milo object
quads_milo <- Milo(sce)

# construct KNN graph
quads_milo <- buildGraph(quads_milo, k = 30, d = 30, 
                            reduced.dim = "HARMONY")

# define representative neighbourhoods
quads_milo <- makeNhoods(quads_milo, 
                                    prop = 0.1, #if ncells<30k use prop=0.1, otherwise use 0.05
                                    # in the paper they use 0.3, and have 90k cells
                            k = 30, d=30, refined = TRUE, reduced_dims = "HARMONY") # paper uses "PCA"
quads_milo
```

Plot the distribution of neighbourhood sizes, we are looking for avg neighbourhood size > 5 X N_samples
Quads data has 6 samples
5 x 6 = 30
Maybe need to adjust k and d to increase the neighbourhood sizes. 

```{r}
plotNhoodSizeHist(quads_milo)

#ggsave(paste0(directory, "/Figures/NeighbourhoodSizeHistogram.png"))
```

Compute neighbourhood connectivity
(NB this step can take a few minutes i.e. half an hour)
```{r}
quads_milo <- calcNhoodDistance(quads_milo, d=30, 
                                   reduced.dim = "HARMONY")
```

Count how many cells are in each neighbourhood per sample
```{r}
quads_milo <- countCells(quads_milo, 
                            meta.data = as.data.frame(colData(quads_milo)),
                            sample="patient")
head(nhoodCounts(quads_milo))
```


## 3. Set up experimental design for differential abundance testing


```{r}
quads_design <- data.frame(colData(quads_milo))[,c("patient", "tendon_disease")]

## Convert batch info from integer to factor
#quads_design$patient <- as.factor(quads_design$patient) 
quads_design <- distinct(quads_design) # keep unique rows

# add rownames (sample names)
rownames(quads_design) <- quads_design$patient

# make sure the design table is in the same order as the nhoodCounts table
quads_design <- quads_design[colnames(nhoodCounts(quads_milo)), , drop=FALSE]
table(quads_design$tendon_disease)# shows how many samples at each microanatomy

quads_design
```


## 4. Perform the differential abundance test  
We want to test for differences between healthy and disease.  
Could account for patient here if required  

```{r}
da_results <- testNhoods(quads_milo, 
                         design = ~tendon_disease,
                         design.df=quads_design, 
                         fdr.weighting="k-distance", # default
                         reduced.dim = "HARMONY"
                         )

da_results$Diff <- sign(da_results$logFC) # add a column to show up or down regulation
da_results$Diff[da_results$SpatialFDR > 0.1] <- 0 # made that column 0 if the result is not significant
table(da_results$Diff)
```

This table shows how many nodes show up- and down-regulation  


## 5. Plot the outputs

#### Inspect the distribution of p values

```{r}
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)
#ggsave(paste0(directory, "/Figures/P-values_distribution.png"))
```

#### Volcano plot of test results
[presumably could use EnhancedVolcano here if preferred]


```{r}
max.lfc <- max(abs(da_results$logFC))

ggplot(da_results, aes(x=logFC, y=-log10(SpatialFDR), colour=as.character(Diff))) +
    geom_hline(yintercept=-log10(0.1), lty=2, colour='grey50') +
    geom_point(size=3) +
    annotate("text", x=-3, y=3, label="Higher in Healthy", size=4) +
    annotate("text", x=3, y=3, label="Higher in Rupture", size=4) +
    scale_x_continuous(limits=c(-max.lfc-0.1, max.lfc+0.1)) +
    theme_cowplot() +
    scale_colour_manual(values=c("firebrick2", "grey80", "firebrick2")) +
    guides(colour="none") +
    labs(x="log Fold Change", y=expression(paste("-log"[10], " Spatial FDR")))

#ggsave(paste0(directory, "/Figures/Volcano_plot.png"), width = 6, height = 6, bg = "white")
```

#### Neighbourhood graph on Harmony UMAP projection 

First build the neighbourhood graph, then plot it


```{r, warning=FALSE, fig.height=5.95, fig.width=6.95}
quads_milo <- buildNhoodGraph(quads_milo, overlap=5)
p1 <- plotNhoodGraphDA(quads_milo, layout="UMAP", milo_res=da_results, alpha=0.1)
p1 
#ggsave(paste0(directory, "/Figures/Neighbourhood_graph_DA.png"), width = 7, height = 6)

```

Do the same plot but colour by original cluster

```{r, fig.width=9, fig.height=6}
p2 <- plotNhoodGraph(quads_milo, layout="UMAP", colour_by="immune_clusterid2") +    scale_fill_manual(values=quads.colours) +
    guides(fill=guide_legend(title="Cell type", override.aes=list(size=4)))
p2
#ggsave(paste0(directory, "/Figures/Neighbourhood_graph_annotation.png"), width = 9, height = 6)

```

Plot both neighbourhood graphs together

```{r, fig.width=14, fig.height=5}

p <- plot_grid (p1, p2, rel_widths = c(0.45, 0.55))
p

#ggsave(paste0(directory, "/Figures/Neighbourhood_graphs.png"), width = 14, height = 5)
```



```{r, fig.width = 10, fig.height=5}
## Plot single-cell UMAP
umap_pl <- plotReducedDim(quads_milo, dimred = "UMAP", 
                          colour_by="tendon_disease", 
                          text_by = "immune_clusterid2", 
                          text_size = 3, point_size=0.5) +
    scale_colour_manual(values = rupture.cols)+
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(quads_milo, da_results,
                                layout="UMAP",alpha=0.1) 
  
p <- plot_grid(umap_pl, nh_graph_pl)
p

#ggsave(paste0(directory, "/Figures/UMAP_vs_Neighbourhood_graph_DA.png"), width = 10, height = 5, bg = "white")
```


#### Annotate neighbourhoods

Annotate the neighbourhoods & plot the homogeneity  
```{r}
da_results <- annotateNhoods(quads_milo, da_results, coldata_col = "immune_clusterid2")
ggplot(da_results, aes(immune_clusterid2_fraction)) + geom_histogram(bins=50)
#ggsave(paste0(directory, "/Figures/Neighbourhood_homogeneity.png"), bg = "white")
```
Annotate as "Mixed" if there are a mixture of cell types

```{r}
da_results$immune_clusterid2 <- ifelse(da_results$immune_clusterid2_fraction < 0.7, "Mixed", da_results$immune_clusterid2)
```

Print the top DA results

```{r}
print ("Top cell types in Rupture")
a <- da_results %>% 
    arrange(abs(SpatialFDR)) %>% 
    filter(Diff == 1) %>%
    select(logFC, SpatialFDR, Diff, cluster_id) %>% 
    head(n=50)
print(a)

print ("Top cell types in Healthy")
b <- da_results %>% 
    arrange(abs(SpatialFDR)) %>% 
    filter(Diff == -1) %>%
    select(logFC, SpatialFDR, Diff, cluster_id) %>% 
    head(n=50)
print (b)

```


#### Beeswarm plot

Visualise the fold changes in different cell types

```{r, fig.height=8, fig.width=8}

da_results_subset <- da_results %>% filter(immune_clusterid2 != "Mixed")

plotDAbeeswarm(da_results, group.by = "immune_clusterid2")+
    theme(axis.title.y=element_blank())
#ggsave(paste0(directory, "/Figures/Beeswarm_plot.png"), width = 10, height = 10, bg = "white")

```

### Volcano plot coloured by cell type

```{r, fig.width=8, fig.height=4}
ggplot(da_results, aes(x=logFC, y=-log10(SpatialFDR), colour=immune_clusterid2)) +
    geom_hline(yintercept=-log10(0.1), lty=2, colour='grey50') +
    geom_point(size=2) +
    annotate("text", x=-4, y=3, label="Higher in Healthy", size=4) +
    annotate("text", x=4, y=3, label="Higher in Rupture", size=4) +
    scale_x_continuous(limits=c(-max.lfc-0.1, max.lfc+0.1)) +
    theme_cowplot() +
    scale_colour_manual(values=quads.colours) +
    #guides(colour="none") +
    labs(x="log Fold Change", y=expression(paste("-log"[10], " Spatial FDR")))

#ggsave(paste0(directory, "/Figures/Volcano_plot_annotated.png"), width = 8, height = 4, bg = "white")
```


#### Save the results  

```{r}
# save the object
saveRDS(quads_milo, "/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS/Quads-immune-milo.rds")

# save the results
write.table(da_results, 
            file = "/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/Quads_analysis_RDS//Quads-immune-milo-DA-results.txt",
            sep = "\t", quote = FALSE, row.names = FALSE)
```


Print session info

```{r}
sessionInfo()
```

