---
title: "MiloR_Cell_proportions-Quads"
author: "Carla Cohen"
date: "`r Sys.Date()`"
output: html_document
---

# Cell type proportion analysis

Aim: To analyse the how cell proportions vary between healthy and tear in quadriceps tendon

Using MioR as recommended by Bo Sun & Lucy Garner

https://www.nature.com/articles/s41587-021-01033-z  
https://marionilab.github.io/miloR/articles/milo_demo.html  
https://www.bioconductor.org/packages/devel/bioc/vignettes/miloR/inst/doc/milo_gastrulation.html  
https://github.com/MarioniLab/miloR  

## Steps

1. Set up 
- Set up directories, import yml file, set colours  
- Read in Seurat object  

2. Set up the milo object & perform initial calculations
- convert so to sce  
- create a milo object  
- construct KNN graph  
- define representative neighbourhoods & plot  
- compute neighbourhood connectivity  
- count number of cells per neighbourhood  

3.  Set up the experimental design for differential abundance testing  

4.  Perform the differential abundance test  

5. Plot the outputs 
- Inspect the distribution of p values  
- Volcano plot of test results
- Neighbourhood graph on Harmony UMAP projection  
- Beeswarm plot  



### 1. Set up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(Seurat)
library(tidyverse)
library(miloR)
library(SingleCellExperiment)
library(scater)
library(scran)
library(yaml)
library(cowplot)

set.seed(1) 

# make a new output folder for each run, with the date & time in the directory name
date <- Sys.Date() %>% str_replace_all("-", "")
time <- format(Sys.time(), "%X") %>% str_replace_all(":", "-") %>%
    str_sub(1, 5)
directory <- paste0(date,"_", time, "_MiloR_Cell_Proportions.dir")
dir.create(directory, showWarnings = FALSE)
dir.create(paste0(directory, "/Figures/"))
dir.create(paste0(directory, "/RDS_objects.dir/"))
dir.create(paste0(directory, "/Results.dir/"))

# read yaml file
ini <- read_yaml("miloR.yml")

# colours
rupture.cols <-  c(Healthy = "#4b2480", Rupture = "#eb7323")

# quads colours
quads.colours <- c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100")

names(quads.colours) <- c("Fibroblasts", "Macrophages", "Vascular endothelial cells",  "Mural cells",  "Adipocytes", "T cells", "Nervous system cells", "Lymphatic endothelial cells", "Dividing fibroblasts / mural cells", "Dendritic cells", "Osteoblasts", "Granulocytes", "Osteoclasts", "Dividing macrophages")
```

Print the parameters used

```{r}
ini
```



Read in Seurat object

```{r}
so.quads <- readRDS("/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20231130_quads_int_labelled.RDS")
so.quads

```

## 2. Set up the milo object & perform initial calculations  
Done following the vignette as I have done in the script miloR_vignette (on computer not cluster)

```{r}

# Convert to single cell experiment object  
sce <- as.SingleCellExperiment(so.quads)

# Create a milo object
quads_milo <- Milo(sce)

# construct KNN graph
quads_milo <- buildGraph(quads_milo, k = 30, d = 30, 
                            reduced.dim = ini$dim_reduction)

# define representative neighbourhoods
quads_milo <- makeNhoods(quads_milo, 
                                    prop = 0.1, #if ncells<30k use prop=0.1, otherwise use 0.05
                                    # in the paper they use 0.3, and have 90k cells
                            k = 30, d=30, refined = TRUE, reduced_dims = ini$dim_reduction) # paper uses "PCA"
quads_milo
```

Plot the distribution of neighbourhood sizes, we are looking for avg neighbourhood size > 5 X N_samples
Quads data has 6 samples
5 x 6 = 30
Maybe need to adjust k and d to increase the neighbourhood sizes. 

```{r}
plotNhoodSizeHist(quads_milo)

ggsave(paste0(directory, "/Figures/NeighbourhoodSizeHistogram.png"))
```

Compute neighbourhood connectivity
(NB this step can take a few minutes i.e. half an hour)
```{r}
quads_milo <- calcNhoodDistance(quads_milo, d=30, 
                                   reduced.dim = ini$dim_reduction)
```

Count how many cells are in each neighbourhood per sample
```{r}
quads_milo <- countCells(quads_milo, 
                            meta.data = as.data.frame(colData(quads_milo)),
                            sample="patient")
head(nhoodCounts(quads_milo))
```


## 3. Set up experimental design for differential abundance testing


```{r}
quads_design <- data.frame(colData(quads_milo))[,c("patient", "tendon_disease")]

## Convert batch info from integer to factor
#quads_design$patient <- as.factor(quads_design$patient) 
quads_design <- distinct(quads_design) # keep unique rows

# add rownames (sample names)
rownames(quads_design) <- quads_design$patient

# make sure the design table is in the same order as the nhoodCounts table
quads_design <- quads_design[colnames(nhoodCounts(quads_milo)), , drop=FALSE]
table(quads_design$tendon_disease)# shows how many samples at each microanatomy

quads_design
```


## 4. Perform the differential abundance test  
We want to test for differences between healthy and disease.  
Could account for patient here if required  

```{r}
da_results <- testNhoods(quads_milo, 
                         design = as.formula(ini$design_formula),
                         design.df=quads_design, 
                         fdr.weighting="k-distance", # default
                         reduced.dim = ini$dim_reduction
                         )

da_results$Diff <- sign(da_results$logFC) # add a column to show up or down regulation
da_results$Diff[da_results$SpatialFDR > 0.1] <- 0 # made that column 0 if the result is not significant
table(da_results$Diff)
```

This table shows how many nodes show up- and down-regulation  


## 5. Plot the outputs

#### Inspect the distribution of p values

```{r}
ggplot(da_results, aes(PValue)) + geom_histogram(bins=50)
ggsave(paste0(directory, "/Figures/P-values_distribution.png"))
```

#### Volcano plot of test results
[presumably could use EnhancedVolcano here if preferred]


```{r, fig.width=6, fig.height=6}
max.lfc <- max(abs(da_results$logFC))

ggplot(da_results, aes(x=logFC, y=-log10(SpatialFDR), colour=as.character(Diff))) +
    geom_hline(yintercept=-log10(0.1), lty=2, colour='grey50') +
    geom_point(size=3) +
    annotate("text", x=-3, y=3, label="Higher in Healthy", size=4) +
    annotate("text", x=3, y=3, label="Higher in Rupture", size=4) +
    scale_x_continuous(limits=c(-max.lfc-0.1, max.lfc+0.1)) +
    theme_cowplot() +
    scale_colour_manual(values=c("#4b2480", "grey80", "#eb7323")) +
    guides(colour="none") +
    labs(x="log Fold Change", y=expression(paste("-log"[10], " Spatial FDR")))

#ggsave(paste0(directory, "/Figures/Volcano_plot.png"), width = 6, height = 6, bg = "white")
```

#### Neighbourhood graph on Harmony UMAP projection 

First build the neighbourhood graph, then plot it


```{r, warning=FALSE, fig.height=8, fig.width=8}
quads_milo <- buildNhoodGraph(quads_milo, overlap=5)
p1 <- plotNhoodGraphDA(quads_milo, layout="UMAP", milo_res=da_results, alpha=0.1) + theme(legend.position = "bottom", legend.box = "vertical")
p1 
ggsave(paste0(directory, "/Figures/Neighbourhood_graph_DA.png"), width = 7, height = 6)

```

Do the same plot but colour by original cluster

```{r, fig.width=8, fig.height=8}
p2 <- plotNhoodGraph(quads_milo, layout="UMAP", colour_by="cluster_id") +    scale_fill_manual(values=quads.colours) +
    guides(fill=guide_legend(title="Cell type", override.aes=list(size=4), ncol = 3)) + theme(legend.position = "bottom", legend.box = "vertical")
p2
ggsave(paste0(directory, "/Figures/Neighbourhood_graph_annotation.png"), width = 9, height = 6)

```

Plot both neighbourhood graphs together

```{r, fig.width=14, fig.height=5}

p <- plot_grid (p1, p2, rel_widths = c(0.45, 0.55))
p

ggsave(paste0(directory, "/Figures/Neighbourhood_graphs.png"), width = 14, height = 5)
```



```{r, fig.width = 10, fig.height=5}
## Plot single-cell UMAP
umap_pl <- plotReducedDim(quads_milo, dimred = "UMAP", 
                          colour_by="tendon_disease", 
                          text_by = "cluster_id", 
                          text_size = 3, point_size=0.5) +
    scale_colour_manual(values = rupture.cols)+
  guides(fill="none")

## Plot neighbourhood graph
nh_graph_pl <- plotNhoodGraphDA(quads_milo, da_results,
                                layout="UMAP",alpha=0.1) 
  
p <- plot_grid(umap_pl, nh_graph_pl)
p

ggsave(paste0(directory, "/Figures/UMAP_vs_Neighbourhood_graph_DA.png"), width = 10, height = 5, bg = "white")
```


#### Annotate neighbourhoods

Annotate the neighbourhoods & plot the homogeneity  
```{r}
da_results <- annotateNhoods(quads_milo, da_results, coldata_col = "cluster_id")
ggplot(da_results, aes(cluster_id_fraction)) + geom_histogram(bins=50)
ggsave(paste0(directory, "/Figures/Neighbourhood_homogeneity.png"), bg = "white")
```
Annotate as "Mixed" if there are a mixture of cell types

```{r}
da_results$cluster_id <- ifelse(da_results$cluster_id_fraction < 0.7, "Mixed", da_results$cluster_id)
```

Print the top DA results

```{r}
print ("Top cell types in Rupture")
a <- da_results %>% 
    arrange(abs(SpatialFDR)) %>% 
    filter(Diff == 1) %>%
    select(logFC, SpatialFDR, Diff, cluster_id) %>% 
    head(n=50)
print(a)

print ("Top cell types in Healthy")
b <- da_results %>% 
    arrange(abs(SpatialFDR)) %>% 
    filter(Diff == -1) %>%
    select(logFC, SpatialFDR, Diff, cluster_id) %>% 
    head(n=50)
print (b)

```


#### Beeswarm plot

Visualise the fold changes in different cell types

```{r, fig.height=9, fig.width=10}

da_results_subset <- da_results %>% filter(cluster_id != "Mixed")

plotDAbeeswarm(da_results_subset, group.by = "cluster_id")+
    scale_colour_gradient2(low = muted("#4b2480"), high = muted("#eb7323")) +
    theme(axis.title.y=element_blank())
ggsave(paste0(directory, "/Figures/Beeswarm_plot.png"), width = 10, height = 10, bg = "white")

```

### Volcano plot coloured by cell type

```{r, fig.width=6, fig.height=7.7}
p3 <- ggplot(da_results, aes(x=logFC, y=-log10(SpatialFDR), colour=cluster_id)) +
    geom_hline(yintercept=-log10(0.1), lty=2, colour='grey50') +
    geom_point(size=1.5) +
    annotate("text", x=-4, y=3, label="Higher in Healthy", size=4) +
    annotate("text", x=4, y=3, label="Higher in Rupture", size=4) +
    scale_x_continuous(limits=c(-max.lfc-0.1, max.lfc+0.1)) +
    theme_cowplot() +
    scale_colour_manual(values=quads.colours) +
    labs(x="log Fold Change", y=expression(paste("-log"[10], " Spatial FDR")))  + theme(legend.position = "bottom", legend.box = "vertical")

p3 + guides(col = guide_legend(title="Cell type", ncol = 2))

ggsave(paste0(directory, "/Figures/Volcano_plot_annotated.png"), width = 8, height = 4, bg = "white")
```


#### Save the results  

```{r}
# save the object
saveRDS(quads_milo, paste0(directory, "/RDS_objects.dir/Quads-milo.rds"))

# save the results
write.table(da_results, 
            file = paste0(directory, "/Results.dir/Quads-milo-DA-results.txt"),
            sep = "\t", quote = FALSE, row.names = FALSE)

```


#If anything needs to be re-run, open files from saved locations:
```{r}
quads_milo <- readRDS("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20240411_11-59_MiloR_Cell_Proportions.dir/RDS_objects.dir/Quads-milo.rds")
da_results <- read.table("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20240411_11-59_MiloR_Cell_Proportions.dir/Results.dir/Quads-milo-DA-results.txt", sep = "\t", header = T)
```


Print session info

```{r}
sessionInfo()
```

