---
title: "Quads_merge_int_cluster_plot.Rmd"
author: "Jolet Mimpen"
date: "2025-01-21"
output: html_document
---


```{r setup, load packages, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(Seurat)
library(patchwork)
library(tidyverse)
library(Matrix)
library(ggplot2)
library(scales)
library(RColorBrewer)
library(harmony)
library(cowplot)
library(clustree)
library(sctransform)
#library(scCustomize)
library(scuttle)
library(celda)

```


#Load in all Seurat files
```{r, read in objects}

# Read in Seurat objects
file_list <- list.files("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20231114_decontX_SoupX_scDblFinder/RDS_objects/")

#Make a new list for the so objects
Seurat.list <- list()
# generate a list of so objects
for (i in 1:length(file_list)){
    print(i)
        Seurat.list[[i]] <- readRDS(paste0("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/20231114_decontX_SoupX_scDblFinder/RDS_objects/", file_list[i]))
    print(file_list[i])
    print(Seurat.list[[i]])
}

```



#Change metadata
```{r}
#sample 1 MSK0778 tear
#sample 2 MSK0779-1 tear
#sample 3 MSK0779-2 tear
#sample 4 MSK0792 healthy
#sample 5 MSK0793 tear
#sample 6 MSK1248 healthy
#sample 7 MSK1250 healthy
#sample 8 MSK1266 healthy

Seurat.list[[1]]$tendon_disease <- "Rupture"
Seurat.list[[2]]$tendon_disease <- "Rupture"
Seurat.list[[3]]$tendon_disease <- "Rupture"
Seurat.list[[4]]$tendon_disease <- "Healthy"
Seurat.list[[5]]$tendon_disease <- "Rupture"
Seurat.list[[6]]$tendon_disease <- "Healthy"
Seurat.list[[7]]$tendon_disease <- "Healthy"
Seurat.list[[8]]$tendon_disease <- "Healthy"

Seurat.list[[6]]$disease_status <- "Tibial shaft fracture"


Seurat.list[[6]]$surgical_procedure <- "Suprapatellar IM nail surgery"
Seurat.list[[8]]$surgical_procedure <- "Suprapatellar IM nail surgery"


#double-check that 7th item in the list is MSK1250
Seurat.list[[7]]$orig.ident[1]

```



#Merge datasets, leaving out the MSK1250 sample
```{r}
Seurat.list.merged <- Reduce(function(x, y){merge(x,y)}, Seurat.list[- 7])

rm(Seurat.list)
```


#Filter out high ambient contamination and doublets
```{r, fig.width=15, fig.height=6}

Seurat.list.merged <- subset(Seurat.list.merged, subset = scDblFinder.class == "singlet" & decontX_contamination < 0.3)


Idents(Seurat.list.merged) <- Seurat.list.merged@meta.data$orig.ident

VlnPlot(Seurat.list.merged, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2, log = T)
VlnPlot(Seurat.list.merged, features = c("nFeature_SoupXcounts", "nCount_SoupXcounts"), ncol = 2, log = T)
VlnPlot(Seurat.list.merged, features = c("subsets_mito_percent", "decontX_contamination"), ncol = 2)
```



#Lognormalise all assays, run PCAs on SoupX assay
```{r}

DefaultAssay(object = Seurat.list.merged) <- "RNA"

data.integratedLOG <- Seurat.list.merged %>% 
                NormalizeData() %>%
                FindVariableFeatures(selection.method = "vst", nfeatures = 5000) %>% 
                ScaleData(verbose=FALSE)

DefaultAssay(object = data.integratedLOG) <- "decontXcounts"

data.integratedLOG <- data.integratedLOG %>% 
                NormalizeData() %>%
                FindVariableFeatures(selection.method = "vst", nfeatures = 5000) %>% 
                ScaleData(verbose=FALSE)

DefaultAssay(object = data.integratedLOG) <- "SoupXcounts"

data.integratedLOG <- data.integratedLOG %>% 
                NormalizeData() %>%
                FindVariableFeatures(selection.method = "vst", nfeatures = 5000) %>% 
                ScaleData(verbose=FALSE) %>%
                RunPCA(npcs = 50, verbose=FALSE)
```


#Run Harmony integration
```{r}
data.integratedLOG.h <- RunHarmony(data.integratedLOG, group.by.vars = "orig.ident", plot_convergence = TRUE)

```



#Make Elbow plots
```{r}
ElbowPlot(data.integratedLOG.h, ndims = 50, reduction = "pca")
ElbowPlot(data.integratedLOG.h, ndims = 50, reduction = "harmony")

```




#Make clustree
```{r, fig.height=8, fig.width=8}
library(clustree)

dims.choice <- 40
resolution.range <- seq(from = 0, to = 0.8, by = 0.1)
data.integratedLOG.h <- FindNeighbors(data.integratedLOG.h, reduction = "harmony", dims = 1:dims.choice)
data.integratedLOG.h <- FindClusters(data.integratedLOG.h, resolution = resolution.range)

clustree(data.integratedLOG.h, prefix = "SoupXcounts_snn_res.")



```



#UMAPs for every resolution
```{r}

data.integratedLOG.h <- RunUMAP(data.integratedLOG.h, reduction = "harmony", dims = 1:dims.choice)

for (i in resolution.range){
  Idents(data.integratedLOG.h) <- paste0("SoupXcounts_snn_res.", i)
  print(DimPlot(data.integratedLOG.h, reduction = "umap", label = T) + ggtitle(i) + theme(plot.title = element_text(hjust = 0.5)))
}

```





#Clustering and plotting
```{r, fig.width=10}

Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.2"


DimPlot(data.integratedLOG.h, reduction = "umap", label = T) +
FeaturePlot(data.integratedLOG.h, reduction = "umap", features = "nCount_RNA")
FeaturePlot(data.integratedLOG.h, reduction = "umap", features = "nFeature_RNA", max.cutoff = 1000)
FeaturePlot(data.integratedLOG.h, reduction = "umap", features = "subsets_mito_percent")
DimPlot(data.integratedLOG.h, reduction = "umap", group.by = "scDblFinder.class")
DimPlot(data.integratedLOG.h, reduction = "umap", group.by = "patient")
DimPlot(data.integratedLOG.h, reduction = "umap", group.by = "disease_status")
DimPlot(data.integratedLOG.h, reduction = "umap", group.by = "tendon_disease")
FeaturePlot(data.integratedLOG.h, reduction = "umap", features = "decontX_contamination", max.cutoff = 0.3)
FeaturePlot(data.integratedLOG.h, reduction = "umap", features = "nCount_SoupXcounts", max.cutoff = 2000)
FeaturePlot(data.integratedLOG.h, reduction = "umap", features = "nFeature_SoupXcounts", max.cutoff = 1000)
VlnPlot(data.integratedLOG.h, features = "nFeature_RNA", log = T)
VlnPlot(data.integratedLOG.h, features = "nCount_RNA", log = T)
VlnPlot(data.integratedLOG.h, features = "subsets_mito_percent")
VlnPlot(data.integratedLOG.h, features = "decontX_contamination")
VlnPlot(data.integratedLOG.h, features = "nCount_SoupXcounts", log = T)
VlnPlot(data.integratedLOG.h, features = "nFeature_SoupXcounts", log = T)

```

#UMAP split by surgical procedure
```{r, fig.width=15, fig.height=6}

DimPlot(data.integratedLOG.h, reduction = "umap", split.by = "surgical_procedure")
DimPlot(data.integratedLOG.h, reduction = "umap", split.by = "tendon_disease")

```



#UMAP per sample
```{r, fig.width=15, fig.height=10}
DimPlot(data.integratedLOG.h, reduction = "umap", split.by = "patient", ncol = 3)
```




#General overview
```{r, fig.width=15, fig.height=5}
#multiple genes find ENS name and create data frame with both names

geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN",
              "NEGR1", "HMCN1", "DCLK1",
              "CLU", "COMP",  "CILP", "NOX4", "PRG4",
              "PTPRC", "CD163", "MRC1", "MSR1", "MERTK", "LYVE1", "CLEC10A",
              "PECAM1", "PTPRB", "FLT1", "VWF",
              "TRDN", "TTN", "NEB", "TNNT1", "TNNT3", "PAX7",
              "NOTCH3", "PDGFRB", "MYO1B",
              "PLIN1", "AQP7", "ADIPOQ", 
              "CD247", "SKAP1", "THEMIS",
              "MMRN1", "PROX1", "PKHD1L1",
              "ASPM", "DIAPH3", "TOP2A",
              "CLEC4C", "CUX2", "IRF7", "IRF8",
              "IBSP", "SP7", "RUNX2",
               "KIT", "CPA3", "IL18R1",
              "MMP9", "ACP5", "SIGLEC15",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2")

geneName <- c("PTPRC",
              "PLIN1", "PLIN4", "AQP7", "ADIPOQ", 
              "CLEC4C", "CUX2", "IRF7", "IRF8",
              "COL1A1", "COL1A2", "COL3A1", "DCN", "NEGR1", "HMCN1", "DCLK1", "NOX4", 
              "ASPM", "DIAPH3", "TOP2A",
              "KIT", "CPA3", "IL18R1",
              "MMRN1", "PROX1", "PKHD1L1",
              "CD163", "MRC1", "MSR1", "MERTK", "LYVE1",
              "NOTCH3", "PDGFRB", "MYO1B",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2",
              "IBSP", "SP7", "RUNX2", "SPP1",
              "MMP9", "ACP5", "SIGLEC15",
              "CD247", "SKAP1", "THEMIS", "IL7R",
              "PECAM1", "PTPRB", "FLT1", "VWF")


DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#Dotplots resolutions 0.1 to 0.6
```{r, fig.width=15, fig.height=5}
geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN",
              "NEGR1", "HMCN1", "DCLK1",
              "CLU", "COMP",  "CILP", "NOX4", "PRG4",
              "PTPRC", "CD163", "MRC1", "MSR1", "MERTK", "LYVE1", "CLEC10A",
              "PECAM1", "PTPRB", "FLT1", "VWF",
              "TRDN", "TTN", "NEB", "TNNT1", "TNNT3", "PAX7",
              "NOTCH3", "PDGFRB", "MYO1B",
              "PLIN1", "AQP7", "ADIPOQ", 
              "CD247", "SKAP1", "THEMIS",
              "MMRN1", "PROX1", "PKHD1L1",
              "ASPM", "DIAPH3", "TOP2A",
              "BCL11A", "CUX2", "CLEC4C",
              "IBSP", "SP7", "RUNX2",
               "KIT", "CPA3", "IL18R1",
              "MMP9", "ACP5", "SIGLEC15",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2")

Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.1"

DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.1")



Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.2"

DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.2")



Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.3"

DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.3")



Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.4"

DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.4")



Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.5"

DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.5")



Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.6"

DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.6")

```


#Find markers for cluster 6
```{r}
cluster6 <- FindMarkers(data.integratedLOG.h, ident.1 = "6", only.pos = T)

cluster6$geneName <- rownames(cluster6)
```


#Subset Cluster6
```{r, fig.width=12, fig.height=4}
#Choose one of the two options below:
so.cluster6 <- subset(data.integratedLOG.h, idents = c("6"))

#Then continue with Seurat/Harmony
so.cluster6 <- so.cluster6 %>% 
                Seurat::NormalizeData(verbose = FALSE) %>%
                FindVariableFeatures(selection.method = "vst", nfeatures = 2000) %>% 
                ScaleData(verbose=FALSE) %>% 
                RunPCA(npcs = 50, verbose=FALSE)

so.cluster6 <- RunHarmony(so.cluster6, group.by.vars = "orig.ident", plot_convergence = TRUE)

dims.choice <- 30
so.cluster6 <- so.cluster6 %>% 
  RunUMAP(reduction = "harmony", dims = 1:dims.choice) %>% 
  FindNeighbors(reduction = "harmony", dims = 1:dims.choice) %>% 
  FindClusters(resolution = 0.3) %>% 
  identity()
DimPlot(so.cluster6, reduction = "umap", label = TRUE, pt.size = 0.7)
DimPlot(so.cluster6, reduction = "umap", split.by = "tendon_disease")

geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN",
              "NEGR1", "HMCN1", "DCLK1",
              "CLU", "COMP",  "CILP", "NOX4", "PRG4",
              "PTPRC", "CD163", "MRC1", "MSR1", "MERTK", "LYVE1", "CLEC10A",
              "PECAM1", "PTPRB", "FLT1", "VWF",
              "TRDN", "TTN", "NEB", "TNNT1", "TNNT3", "PAX7",
              "NOTCH3", "PDGFRB", "MYO1B",
              "PLIN1", "AQP7", "ADIPOQ", 
              "CD247", "SKAP1", "THEMIS",
              "MMRN1", "PROX1", "PKHD1L1",
              "ASPM", "DIAPH3", "TOP2A",
              "BCL11A", "CUX2", "CLEC4C",
              "IBSP", "SP7", "RUNX2",
               "KIT", "CPA3", "IL18R1",
              "MMP9", "ACP5", "SIGLEC15",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2")


DotPlot(so.cluster6, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - cluster 6 only - res0.3")



```




#Make Heatmap cluster 6 only
```{r, fig.width=10, fig.height=10}
cluster6.submarkers <- FindAllMarkers(so.cluster6, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

cluster6.submarkers %>%
    group_by(cluster) %>%
    top_n(n = 25, wt = avg_log2FC) -> cluster6.submarkers.top25


DoHeatmap(so.cluster6, features = cluster6.submarkers.top25$gene) + scale_fill_viridis() + NoLegend()
```

#Add labels for each cluster within Cluster6 object
```{r, fig.height=6, fig.width=10}


new.cluster.ids <- c("Nervous system cells", "Dendritic cells") 


names(new.cluster.ids) <- levels(so.cluster6)
so.cluster6 <- RenameIdents(so.cluster6, new.cluster.ids)


DimPlot(so.cluster6, reduction = "umap", repel = TRUE, label = F, label.size = 5.0, pt.size = 0.8) + NoLegend()

DimPlot(so.cluster6, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, pt.size = 0.8) + NoLegend()

so.cluster6$clusterid <- Idents(so.cluster6)


#data.integratedLOG.h@meta.data$clusternames_res0.4_long <- Idents(data.integratedLOG.h)

```





#Make Heatmap
```{r, fig.width=15, fig.height=25}
general.markers <- FindAllMarkers(data.integratedLOG.h, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

general.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10


DoHeatmap(data.integratedLOG.h, features = top10$gene) + scale_fill_viridis() + NoLegend()
```



#Fibroblast markers
```{r, fig.width=12, fig.height=5}


#geneName <- c("COL1A1", "COL1A2", "COL3A1",  "DCN", "PRG4", "ABCA8", "COL15A1", "CD55", "DCLK1", "ELN", "FBN1", "FBLN1", "FBLN2", "FBLN5", "KCND2", "NEGR1", "NOVA1", "PDGFRA", "VIT", "CADM1", "COL11A1", "COL11A2", "COL12A1", "COL24A1", "COMP", "CPXM2", "FMOD", "MET", "MKX", "ITGA10", "PIEZO2", "THBS4", "THSD4", "TNMD")

geneName <- c("NOX4", "ITGA11", "DCN", "KCNMA1",  "GALNT15", "KLHL29", "THBS4", "NOVA1", "COMP", "COL14A1",
              "PRG4", "FN1", "ITGB8", "ZNF385B", "CRTAC1", "ITGBL1", "TMEM196", "SEMA5A", "DELEC1", "CLU", 
              "NEGR1", "ABCA10", "PTGFR", "ABCA8", "KAZN", "DCLK1", "ENSG00000231606", "VCAN", "LAMA2", "VCAN-AS1", 
              "SLC35F4", "FGFR2", "POSTN", "ENSG00000231873", "KIF26B", "ADAM12", "NAV3", "GLIS3")


plot5 <- DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"))

plot5 +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
#remove(geneName, plot4) #cleanup
```



#All collagens
```{r, fig.width=10, fig.height=4}

geneName <- c("COL1A1", "COL1A2", "COL2A1", "COL3A1", "COL4A1","COL4A2","COL4A3", "COL4A4", "COL4A5", "COL4A6", "COL5A1", "COL5A2", "COL5A3", "COL6A1", "COL6A2", "COL6A3", "COL6A5", "COL7A1", "COL8A1", "COL8A2", "COL9A1", "COL9A2", "COL9A3", "COL10A1",  "COL11A1", "COL11A2", "COL12A1", "COL13A1", "COL14A1", "COL15A1", "COL16A1", "COL17A1", "COL18A1", "COL19A1", "COL20A1", "COL21A1", "COL22A1", "COL23A1", "COL24A1", "COL25A1", "EMID2", "COL27A1", "COL28A1", "COL29A1")

plot4 <- DotPlot(data.integratedLOG.h, features = geneName, cols = c("blue", "red"))

plot4 +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

```



#Cell cycle markers
```{r, fig.width=14, fig.height=5}
s.genes <- cc.genes$s.genes
g2m.genes <- cc.genes$g2m.genes

Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.1"

dotplot.sgenes.res0.1 <- DotPlot(data.integratedLOG.h, features = s.genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.1 - S genes")
dotplot.sgenes.res0.1

dotplot.g2mgenes.res0.1 <- DotPlot(data.integratedLOG.h, features = g2m.genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.1 - G2M genes")
dotplot.g2mgenes.res0.1




Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.2"

dotplot.sgenes.res0.2 <- DotPlot(data.integratedLOG.h, features = s.genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.2 - S genes")
dotplot.sgenes.res0.2

dotplot.g2mgenes.res0.2 <- DotPlot(data.integratedLOG.h, features = g2m.genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - res0.2 - G2M genes")
dotplot.g2mgenes.res0.2


DotPlot(data.integratedLOG.h, features = s.genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("S genes")


DotPlot(data.integratedLOG.h, features = g2m.genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("G2M genes")



```



```{r, fig.width=12, fig.height=10}
dotplot.cellcyclegenes.res0.1 <- cowplot::plot_grid(dotplot.sgenes.res0.1, dotplot.g2mgenes.res0.1, ncol = 1)

dotplot.cellcyclegenes.res0.1

dotplot.cellcyclegenes.res0.2 <- cowplot::plot_grid(dotplot.sgenes.res0.2, dotplot.g2mgenes.res0.2, ncol = 1)

dotplot.cellcyclegenes.res0.2


```




#Add labels for each cluster of overall object
```{r, fig.height=6, fig.width=10}

Idents(data.integratedLOG.h) <- "SoupXcounts_snn_res.0.2"

new.cluster.ids <- c("Fibroblasts", "Macrophages", "Vascular endothelial cells",  "Mural cells",  "Adipocytes", "T cells", "6", "Lymphatic endothelial cells", "Dividing fibroblasts / mural cells", "Osteoblasts", "Granulocytes", "Osteoclasts", "Dividing macrophages")


names(new.cluster.ids) <- levels(data.integratedLOG.h)
data.integratedLOG.h <- RenameIdents(data.integratedLOG.h, new.cluster.ids)


DimPlot(data.integratedLOG.h, reduction = "umap", repel = TRUE, label = F, label.size = 5.0, pt.size = 0.8) + NoLegend()

DimPlot(data.integratedLOG.h, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, pt.size = 0.8) + NoLegend()

#add names into the metadata
data.integratedLOG.h@meta.data$cluster_id <- Idents(data.integratedLOG.h)

```


#Overwrite names of cells in cluster 6
```{r}
df <- so.cluster6@active.ident

df <- as.data.frame(df)

colnames(df) <- c("clusterid")

data.integratedLOG.h$cluster_idbackup <- data.integratedLOG.h$cluster_id
#data.integratedLOG.h$cluster_id <- data.integratedLOG.h$cluster_idbackup 

df$clusterid <- as.character(df$clusterid)
data.integratedLOG.h$cluster_id <- as.character(data.integratedLOG.h$cluster_id)

for (i in seq_len(nrow(df))){
    barcode <- rownames(df)[i]
    cell_type <- df$clusterid[i]
   
    if (data.integratedLOG.h$cluster_id[barcode] == "6"){
        if (cell_type == "Dendritic cells") {
            data.integratedLOG.h$cluster_id[barcode] <- "Dendritic cells"
        } else if (cell_type == "Nervous system cells") {
            data.integratedLOG.h$cluster_id[barcode] <- "Nervous system cells"
        }
    }
}

Idents(data.integratedLOG.h) <- "cluster_id"
DimPlot(data.integratedLOG.h, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, pt.size = 0.8)

#make new object with new names of all (sub)clusters
quads_data <- data.integratedLOG.h

```


#Re-order the levels of cell types
```{r, fig.width=12, fig.height=6}
quads_data$cluster_id <- factor(quads_data$cluster_id, levels = c("Fibroblasts", "Macrophages", "Vascular endothelial cells",  "Mural cells",  "Adipocytes", "T cells", "Nervous system cells", "Lymphatic endothelial cells", "Dividing fibroblasts / mural cells", "Dendritic cells", "Osteoblasts", "Granulocytes", "Osteoclasts", "Dividing macrophages"))

Idents(quads_data) <- "cluster_id"
```

#UMAPs labelled and unlablled with new colours
```{r, fig.width=9, fig.height=6}

umap_labelled <- DimPlot(quads_data, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, cols = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100"))  + NoLegend()
umap_labelled

umap_notlabelled <- DimPlot(quads_data, reduction = "umap", repel = TRUE, label = F, label.size = 5.0, cols = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100")) + NoLegend()
umap_notlabelled

ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/UMAP_labelled.png", plot = umap_labelled, width = 9, height = 6, bg = "white")
ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/UMAP_notlabelled.png", plot = umap_notlabelled, width = 9, height = 6, bg = "white")

rm(umap_labelled, umap_notlabelled)
```


```{r, fig.width=10, fig.height=5}
genes <- c("NOTCH3", "MYH11", "MYH9", "DES", "MCAM", "RGS5", "ACTA2", "PDGFRB", "COL4A1", "KCNJ8", "MYL9", "ABCC9")

quads_data <- SetIdent(quads_data, value = quads_data$cluster_id)

DotPlot(quads_data, features = genes, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - labelled")

```



#Re-run dotplot
```{r, fig.width=17, fig.height=5}
#for res0.4
geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN",
              "NEGR1", "HMCN1", "DCLK1",
              "CLU", "COMP",  "CILP", "NOX4", "PRG4",
              "PTPRC", "F13A1", "CD163", "MRC1", "MSR1",
              "PECAM1", "PTPRB", "FLT1", "VWF",
              "TRDN", "TTN", "NEB", "TNNT1", "TNNT3", "PAX7",
              "NOTCH3", "PDGFRB", "MYO1B",
              "PLIN1", "AQP7", "ADIPOQ", 
              "CD247", "SKAP1", "THEMIS",
              "MMRN1", "PROX1", "PKHD1L1",
              "ASPM", "DIAPH3", "TOP2A",
              "BCL11A", "CUX2", "CLEC4C",
              "IBSP", "SP7", "RUNX2",
               "KIT", "CPA3", "IL18R1",
              "MMP9", "ACP9", "ACP5", "SIGLEC15",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2")

geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN", "ABI3BP", "NOVA1",  "NEGR1", "HMCN1", "DCLK1", "FBLN1", #fibroblasts
              "PTPRC", "F13A1", "CD163", "MRC1", "MSR1", "MERTK", "FOLR2", #macrophages
              "PECAM1", "PTPRB", "FLT1", "VWF", "EMCN", #vascularendothelialcells
              "NOTCH3", "PDGFRB", "MYO1B", "EBF2",   #mural cells
              "PLIN1", "PLIN4", "AQP7", "ADIPOQ", "LPL",
              "CD247", "SKAP1", "THEMIS", "IL7R", "ANK3",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2",
              "MMRN1", "PROX1", "PKHD1L1", "LYVE1", "SEMA3A",
              "ASPM", "DIAPH3", "TOP2A", "CENPE", "MELK",
              "CLEC4C", "BCL11A", "CUX2", "IRF8", "PLAC8",
              "IBSP", "SP7", "RUNX2",  "ALPL", "GJA1",
              "KIT", "CPA3", "IL18R1", "MS4A2", "GATA2",
              "SPP1", "MMP9", "ACP5", "SIGLEC15", "CA2")

DotPlot(quads_data, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


#use this to plot and change the title for the gene name

quads_data <- SetIdent(quads_data, value = quads_data$cluster_id)

dotplot_overview <- DotPlot(quads_data, features = geneName, cols = c("blue", "red"), assay = "SoupXcounts") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + ggtitle("Quads merged/integrated - SoupX counts - labelled")
dotplot_overview

ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/dotplot_overview.png", plot = dotplot_overview, width = 15, height = 5, bg = "white")

rm(dotplot_overview)
```

#By disease
```{r, fig.width=17, fig.height=8}
geneName <- c("COL1A1", "COL1A2", "COL3A1", "DCN", "ABI3BP", "NOVA1",  "NEGR1", "HMCN1", "DCLK1", "FBLN1", #fibroblasts
              "PTPRC", "F13A1", "CD163", "MRC1", "MSR1", "MERTK", "FOLR2", #macrophages
              "PECAM1", "PTPRB", "FLT1", "VWF", "EMCN", #vascularendothelialcells
              "NOTCH3", "PDGFRB", "MYO1B", "EBF2",   #mural cells
              "PLIN1", "PLIN4", "AQP7", "ADIPOQ", "LPL",
              "CD247", "SKAP1", "THEMIS", "IL7R", "ANK3",
              "IL1RAPL2", "XKR4", "NRXN1", "CADM2",
              "MMRN1", "PROX1", "PKHD1L1", "LYVE1", "SEMA3A",
              "ASPM", "DIAPH3", "TOP2A", "CENPE", "MELK",
              "CLEC4C", "BCL11A", "CUX2", "IRF8", "PLAC8",
              "IBSP", "SP7", "RUNX2",  "ALPL", "GJA1",
              "KIT", "CPA3", "IL18R1", "MS4A2", "GATA2",
              "SPP1", "MMP9", "ACP5", "SIGLEC15", "CA2")

DotPlot(quads_data, features = geneName, cols = c("#4b2480",  "#eb7323"), assay = "SoupXcounts", split.by = "tendon_disease") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))


```


#Heatmap top5
```{r, fig.height=12, fig.width=8}
Idents(quads_data) <- "cluster_idbackup"
DimPlot(quads_data, reduction = "umap", repel = TRUE, label = TRUE, label.size = 5.0, cols = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100"))  + NoLegend()

general.markers <- FindAllMarkers(quads_data, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)


general.markers %>%
    group_by(cluster) %>%
    top_n(n = 5, wt = avg_log2FC) -> top5

top5heatmap <- DoHeatmap(quads_data, features = top5$gene, label = F, group.colors = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100")) + scale_fill_viridis() + NoLegend()
top5heatmap

general.markers %>%
    group_by(cluster) %>%
    top_n(n = 6, wt = avg_log2FC) -> top6

top5heatmap <- DoHeatmap(quads_data, features = top6$gene, label = F, group.colors = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100")) + scale_fill_viridis() + NoLegend()
top5heatmap


ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/Heatmap_top5.png", plot = top5heatmap, width = 8, height = 9, bg = "white")

rm(top5heatmap)

```


#Re-order the levels of patients
```{r, fig.width=12, fig.height=6}
quads_data$patient <- factor(quads_data$patient, levels = c("MSK0792", "MSK1248",  "MSK1266", "MSK0778", "MSK0779", "MSK0793"))


DimPlot(quads_data, reduction = "umap", repel = TRUE, label = F, split.by = "patient", ncol = 3, cols = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100")) + NoLegend()


ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/UMAP_patients.png", plot = umap_patients, width = 12, height = 8, bg = "white")
rm(umap_patients)

```


#UMAP split by tendon disease
```{r, fig.width=15, fig.height=6}
Idents(quads_data) <- "cluster_id"

umap_healthyvsdisease <- DimPlot(quads_data, reduction = "umap", split.by = "tendon_disease", cols = c("#e25822",  "#499999", "#f28e2b",  "#4e79a7",  "#f3c300", "#e68fac", "#8db600",  "#888888",  "#a1caf1", "#be0032", "#117733", "#332288", "#A55194", "#661100"), label = F) + NoLegend() 
umap_healthyvsdisease

ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/UMAP_healthyvsdisease_2.png", plot = umap_healthyvsdisease, width = 15, height = 6, bg = "white")
rm(umap_healthyvsdisease)

```

#UMAP coloured by tendon disease
```{r, fig.width=9, fig.height=6}
Idents(quads_data) <- "cluster_id"

umap_tendondisease <- DimPlot(quads_data, reduction = "umap", group.by = "tendon_disease", cols = c("#4b2480",  "#eb7323"))
umap_tendondisease

DimPlot(quads_data, reduction = "umap", group.by = "tendon_disease", cols = c("#4b2480", "#eb7323")) + theme(legend.position = "bottom")


ggsave("/ceph/project/tendonhca/shared/chromium/analysis/20230922_quadriceps/202311_figures/UMAP_colouredbytendondisease.png", plot = umap_tendondisease, width = 9, height = 6, bg = "white")
rm(umap_tendondisease)

```



